Molecular dynamics simulation of synchronization in driven particles

### 2020-12-28 manuscript/manuscript_ajp_2020.tex
cleaned up latex comments

\begin{abstract}
  %What problem did you study and why is it important?
  We discuss a
  numerical model
  of particles
  confined to a narrow channel 
  driven
  across a washboard potential energy landscape.
  The model 
  exhibits
  synchronization effects that provide insight
  to the behavior of 
  experimental systems 
  %interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields.
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  We present the basics
  of the molecular dynamics simulations
  for particles moving through
  a viscous liquid
  applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to 
  visualize
  the system with
  animations of the particle motion 
  and detect synchronization using
  measurements of
  the particle velocity as a function of applied driving force.
  We include sample code
  and exercises for students
  %that include
  with 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to interaction or an external stimulus \cite{Pikovsky2003}.
Synchronous behaviors where oscillators move in-phase (i.e. in time)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation such as 
birds coordinating wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternating croaking patterns \cite{Aihara2014},
and humans synchronizing clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two pendula on wall-mounted clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986}.
%Dynamical mode locking
The phenomena 
is often studied with simplified computational models
and can be visualized with %demonstrated
with phase plots or Lissajous figures.
A mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
({\it Could these for Fig. 1})

External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance 
%pushing a child on a swing at the high point on her path, or
using pacemaker (an electrical generator)
pulses to regulate a heart beat
or using a pulsed light to control the pattern of flashing
fireflies \cite{Agrawal2013}.
These dynamical drive studies can be performed
on coupled oscillators %\cite{Kuramoto193?}
or a single particle driven across a periodic potential landscape.
A single particle particle oscillator
within a potential well is 
much like a skateboarder in a half-pipe.  % \cite{}.
%Single particle models avoid many-body effects
%while offering insight.
The applied driving force can be constant (dc) or sinusoidal (ac).
With a pure 
dc drive the landscape modulates 
the average particle velocity.
An ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

@@

%
%
%mini review of synchronization in electronics
%careful - the following is all simulations
%Recent studies demonstrate
%that coupled electronic oscillators can be
%programmed for use as microcontrollers for robotic locomotion \cite{Dutta2019}.

Experiments of colloid particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking,
which cannot be easily imaged in electronics.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a useful medium for studying complex dynamical behaviors.  %cite
Colloid particles are micron-sized,
a relatively large size to control and image.
The interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles.
Because colloids are easy to control and study, 
they are used as experimental models
for systems relatively hard to access and visualize,
such as cold atoms or electron gases \cite{Grier2003}. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Mode-locking 
in colloids has been acheived in systems interacting via magnetic dipoles 
or electric forces 
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.

%-----------------------------------------------------------
%particle interaction types
%Particles which interact over long distances
%include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [mor%e detail and references]
%Particles which interact over short distances include
%bubble arrays/emulsions [more systems and references].

%-----------------------------------------------
%environments
Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining landscape 
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as charge carries in quantum wires \cite{Tarucha1995} and
slime molds in microchannels \cite{Gholami2015}.
%Biological systems such as
%neuron axons and capillaries can also be studied
%with these models [more detail and references].
%Many such systems execute local oscillations
%about stable points [elaborate].
%----------------------------------------------------------
%The presence of a modulating surface
%can modify these patterns in a variety of ways,
%changing the onset of dynamical flows,
%and the overall flow patterns.
%
%external forces
%The
%dynamics of particles subject to 
%an applied external force apply to many physical systems.
%For instance the flow of charges in a conductor,
%or cells responding to a chemical gradient.
%The external force 
%increases the diversity of dynamical behaviors,
%and can cause particles to flow in
%a variety of non-linear complex patterns
%that may be 
%synchronized.
%Disordered chaotic dynamics are also possible,
%where irregular, unpredictable time evolution of
%nonlinear systems occurs in mechanical oscillators \cite{chaos}.
%-------------------------------------------------------------

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we focus numerical studies 
of the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
Numerical modeling of colloids can provide mechanistic insight
that can be difficult to achieve in experimental conditions
where Brownian motion and other sources of noise dominate.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can
create Shapiro steps in the force-velocity relationship.

@@

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the landscape and dynamics in this section
and supplementary material.
We summarize
our results 
in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

%Here we model the microscopic dynamics of phase locking in a colloidal model system using molecular dynamics (MD).

\section{Molecular Dynamics Model}
\label{sec:MD}
We use a classical two-dimensional model for 
studying the dynamics of $N$ interacting particles. 
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$.
Particles are subject to
periodic boundary conditions
such that a particle leaving the edges of the system is mapped
back to a position within the simulation 
by $x_i+L=x_i$ and $y_i+L=y_i$.
The units of the simulated variables are in Table ~\ref{tab:1}.

\begin{table}[h!]
\centering
\caption{Simulation units.}
\begin{ruledtabular}
\begin{tabular}{c c p{5cm}}
Quantity & Unit \\
\hline
length &  $a_0$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
dimensionless interaction strength & $q$ \\
effective colloidal charge & $Z^*$ \\
solvent dielectic constant & $\epsilon \epsilon_0$\\
force & $E_0 / a_0$\\
viscosity/damping constant & $\eta$ \\
time &  $\eta / E_0$ \\
velocity &  $ E_0 / \eta a_0$ \\
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}

We model particle interaction forces
$\vec{F}_{ij} = -\nabla U_{ij}(r_{ij})$ 
with
the Yukawa potential energy 
\begin{equation}
  U_{ij}(r_{ij}) = \frac{E_0}{r_{ij}} e^{-\kappa r_{ij}},
\end{equation}
where particle $i$ and $j$ are distance
$r_{ij} = |\vec{r}_i - \vec{r}_j|$ apart.
This 
screened Coulomb potential
%$E_0=2$ scales strength of repulsion
is scaled in terms of energy unit $E_0$
defined in Table ~\ref{tab:1}.
%$ = k q_1 q_2$. %[CHECK scaling/units].
$\kappa = 1/R_0$ is the screening parameter 
that describes the lengthscale at
which particles interact.
We fix $R_0$ to be $a_0$.
In experiments charge screening is observed
due to ions in the suspending fluid and
the charges of surrounding particles
which
reduces the interaction range of individual particles. % \cite{}.
Because the particles interact over short ranges, 
the numerical models can be run efficiency
using a neighbor list algorithm
determined using a cell method.
[explain!]

Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
\vec{F}_{D}(t) = [F_{DC} + F_{AC} \sin(\omega t)] \hat{y},
\end{equation}
with modifiable parameters including
a constant component $F_{DC}$,
and a time dependent component with amplitude $F_{AC}$
and frequency $\omega = 2 \pi f$.
%The frequency is scaled in time units,
%where [finish].

We use several model enviroments to confine the particles,
assuming the confining force arises from a potential function 
 %creating a landscape of potential minima and maxima
 %that modify the force on a particle as a function of position
 $\vec{F}_{l}(\vec{r}) = -\nabla V_l(\vec{r}).$
 The landscape potential $V(\vec{r}) $ are static 
 with fixed minima and maxima
 that are periodic or quasi-periodic,
 as described in Sec.~\ref{sec:results}.
% 
% One example of a periodic landscape is
%% %that used in Sec.~\ref{sec:one}
% where 
% \begin{equation}
%   V_l(y) = V_0 \cos{(N_p \pi y / L)}
% \end{equation}
% where $N_p$ are the number of periods,
% and $V_0$ is an adjustable parameter
% to set the depth of the troughs.
%
 In multi-particle simulations, we confine
 the particles along the $x-$direction 
 using a periodic function 
 \begin{equation}
   U_{q1D}(x) = U_0 \cos{(\pi x / L)}
 \end{equation}.
 %%We could also use a Gaussian function.
 This 
 quasi-one-dimensional geometry
 confines the particles
 primarily to move 
 along the y-direction
 but allows for some lateral motion of particles.
 Otherwise the repulsive interaction between
 particles would cause them to spread throughout the system.

The simulation is controlled by a $for()$ loop
which runs from an initial to maximum $time$ integer.
Each integer timestep
represents a simulation time of $\Delta t=0.001$.
At each timestep
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position
$\vec{r}_i(t+\Delta t)$.
%
The integration is simple
because 
we model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that 
The damping comes from a viscous fluid model
providing a nonconservative force,
modeled as a linear friction $\vec{F}_{drag} = -\eta \vec{v}_i$
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity of the suspending medium is high.
% [finish... reference...]
This model should be familiar to readers
modeling Milliken's oil drops
in a standard classical mechanics text \cite{Taylor2005}.
%This model is encountered in standard texts,
%for instance Ch. 2 of Taylor's Classical Mechanics.

A particle has the equation of motion 
\begin{equation}
\eta \vec{v}_i = \vec{F}_{l,i} + \sum_{i \neq j}^{N} \vec{F}_{ij} + \vec{F}_{D}(t).
\end{equation}
where $\eta = 1$.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location in the simulation
with respect to other particles $j$.
Since the acceleration as zero,
the Verlet method simplifies to 
the Euler method,
which is used to calculate the position at subsequent time steps.
%Most undergraduate students learn the Euler method in Computational Physics and Differential Equations, but don't encounter the Verlet method.
\begin{equation}
\vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t).
\end{equation}

\section{Results}
\label{sec:results}
 We demonstrate how a single particle (Sec.~\ref{sec:one})
 and many particles
 move (Sec.~\ref{sec:sync})
 in response to this applied force in a variety of environments.
 In Sec.~\ref{sec:kink} we set $F_{DC}$
 to zero and track the motion
 of a high density area of a particle chain
 (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.
%
\subsection{Single particle system}
% In subsection headings, only the first word is capitalized.
\label{sec:one}

>>>>

\begin{abstract}
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  exhibit a variety of synchronization effects
  in
  experimental systems 
  %interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields.
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to 
  visualize
  the system with
  numerical representations of the landscape
  and
  animations of the particle motion.
  Our results show a variety of systems that
  synchronize 
  %and detect synchronization using
  %measurements of
  particle velocity as a function of applied driving force,
  and characterize these behaviors
  with force-velocity curves and phase diagrams.
  We include sample code
  and exercises for students
  %that include
  with 
  opportunities
  to reproduce our results and propose
  new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to interaction or an external stimulus \cite{Pikovsky2003}.
Synchronous behaviors where oscillators move in-phase (i.e. in time)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronization such as 
birds coordinating wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternating croaking patterns \cite{Aihara2014},
and humans synchronizing clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two wall-mounted pendulum clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986}.
%Dynamical mode locking
The phenomena 
is often studied with simplified computational models
and can be visualized with %demonstrated
with phase plots or Lissajous figures.
A mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
({\it As we will demonstrate in phase plots in Figs. X, Y, Z})

External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance a
%pushing a child on a swing at the high point on her path, or
pacemaker (an electrical generator)
pulses to regulate a heart beat
or controlling the pattern of flashing
fireflies with a pulsed light \cite{Agrawal2013}.
These behaviors can be modeled with
numerical studies %can be performed
where
a driving force is applied to simple harmonic system,
either
coupled oscillators %\cite{Kuramoto193?}
or a single particle driven across a periodic potential landscape.
A single particle particle oscillator
within a potential well is 
much like a skateboarder in a half-pipe or
a child on a swing
moving back and forth through equilibrium.  % \cite{}.
Single particle studies 
are useful to understand synchronization
in the absence of collective effects.
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold below
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena known as depinning \cite{Reichhardt2017},
and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

@@

%
%
%mini review of synchronization in electronics
%careful - the following is all simulations
%Recent studies demonstrate
%that coupled electronic oscillators can be
%programmed for use as micro-controllers for robotic locomotion \cite{Dutta2019}.

\section{Colloidal Models}
\label{sec:colloids}

Experiments with colloid particles,
such as plastic beads suspended in
highly de-ionized water \cite{},
are useful model systems
to understand complex behaviors.
%silica spheres in organic solvents
%magnetic spheres?
%
Experiments of colloid particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of individual particles during mode-locking,
which cannot be easily imaged in electronics.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a useful medium for studying complex dynamical behaviors  %cite
due to their 
relatively large size of microns,
making them relatively simple 
to control and image.
Additionally colloids can mimic different
systems since 
the interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the particles.
Because colloids are easy to control and study, 
they are used as experimental models
for systems relatively hard to access and visualize,
such as cold atoms or electron gases \cite{Grier2003}. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Mode-locking 
in colloids has been achieved in systems interacting via magnetic dipoles 
or electric forces 
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.

%-----------------------------------------------------------
%particle interaction types
%Particles which interact over long distances
%include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more%e detail and references]
%Particles which interact over short distances include
%bubble arrays/emulsions [more systems and references].----------------------
%The presence of a modulating surface
%can modify these patterns in a variety of ways,
%changing the onset of dynamical flows,
%and the overall flow patterns.
%  adjustable strength, geometry and length scale of substrate potential
%Baumgartl, Brunner, Bechinger, PRL 93, 168301 (2004).
%Baumgartl, Zvyagolskaya, Bechinger, PRL 99, 205503 (2007).

%external forces
%The
%dynamics of particles subject to 
%an applied external force apply to many physical systems.
%For instance the flow of charges in a conductor,
%or cells responding to a chemical gradient.
%The external force 
%increases the diversity of dynamical behaviors,
%and can cause particles to flow in
%a variety of non-linear complex patterns
%that may be 
%synchronized.
%Disordered chaotic dynamics are also possible,
%where irregular, unpredictable time evolution of
%nonlinear systems occurs in mechanical oscillators \cite{chaos}.
%-------------------------------------------------------------

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
Numerical modeling of colloids can provide mechanistic insight
that can be difficult to achieve in experimental conditions
where Brownian motion and other sources of noise dominate.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can
create Shapiro steps in the force-velocity relationship.

@@

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model in Section~\ref{sec:MD}
In this section, we restrict our discussion to a single particle,
and include additional information
for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
We include code to simulate
and visualize the landscape and dynamics in this section
and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

%Here we model the microscopic dynamics of phase locking in a colloidal model syestem using molecular dynamics (MD).

\section{Molecular Dynamics Simulation}
\label{sec:MD}
In this section we describe the dynamics of the particles
using the model forces
and the motion using an modified form of Newton's second law
suitable for microscopic particles in a suspending fluid.

We use a classical two-dimensional model for 
studying the dynamics of $N$ interacting particles.
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$
and velocity $\vec{v}_i = d\vec{r}_i/dt$.  
Particles are subject to
periodic boundary conditions
such that a particle leaving the edge of the system is mapped
back to a position within the simulation 
by the transformation $x_i+L \rightarrow x_i$ and $y_i+L \rightarrow y_i$.
%and particles interact across PBC.

Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{DC} + F^{AC} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{DC}$,
and a time dependent component with amplitude $F^{AC}$
and frequency $\omega = 2 \pi f$.
%The frequency is scaled in time units,
%where [finish].
When $F^{DC}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{DC}$ from zero
to avoid transient effects.

We confine the particles using a position dependent 
potential energy function, called a landscape or substrate.
The confining force %arises from a potential function 
 %creating a landscape of potential minima and maxima
%that modify the force on a particle as a function of position
is calculated as 
 $\vec{F}^{l}_i(\vec{r}_i) = -\nabla U_l(\vec{r}_i)$.
 %The landscape potential energy $U(\vec{r})$ is static 
 %with fixed minima and maxima
 %that are periodic or quasi-periodic,
 %as described in Sec.~\ref{sec:results}.
% 
The periodic landscape is modulated in the y-direction as
%% %that used in Sec.~\ref{sec:one}
% where 
 \begin{equation}
   U_{washboard}(y) = U_0 \cos{(N_p \pi y / L)}
     \label{eq:ysubstrate}
\end{equation}
 where $N_p$ are the number of periods,
 % $L$ is the period length,
 and $U_0$ is an adjustable parameter
 to set the depth of the troughs.
 We define the substrate strength to be
 the force 
 $A_p = 2\pi U_0/L$.
 %

%because 
We model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that
The term ``overdamped'' or critically damped indicates
that an oscillator cannot undergo periodic
undulations due to some external force. %such as
%suspension in a viscous fluid \cite{}.
In our colloidal model we assume
the damping comes from the viscosity of the fluid,
where a sphere of radius $0.5 \mu m$ has a Reynolds number
of $R \approx 10^{-6}$, indicating that linear drag dominates its motion.
See Section~\ref{sec:problems} Exercise~\ref{ex:reynolds} for more information
about the Reynold's number.
We model this
nonconservative force,
with the linear friction $\vec{F}_{drag} = -\eta \vec{v}_i$
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity of the suspending medium is high.
% [finish... discussion of Reynolds number]
This model should be familiar to readers
modeling Millikan's oil drops
in a standard classical mechanics text \cite{Taylor2005}.
%This model is encountered in standard texts,
%for instance Ch. 2 of Taylor's Classical Mechanics.
%more precise: 6\pi \eta a \vec{v}

Newton's second law for an individual particle
$m \vec{a}_i = \sum \vec{F}$ is simplified
by the assumption $\vec{a}_i$ is zero in overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The equation of motion for an isolated particle 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i} + \vec{F}^{D}(t).
    \label{eq:motion}
\end{equation}
where the viscosity coefficient $\eta = 1$.
In Section~\ref{sec:problems} Exercise~\ref{ex:n2l}
we invite the reader to confirm this result.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location $\vec{r}_i$ %in the simulation
with respect to other particles $j$.

The molecular dynamics simulation is controlled by a $for()$ loop
 which runs from an initial to maximum integer number of time steps $t_i$.
Each integer time step interval $t_{i+1}-t_i$
represents a simulation time interval of $\Delta t$ 
simulation units.  %as described in Table 1
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position
$\vec{r}_i(t+\Delta t)$.
%
Since the acceleration as zero,
[more background here - do the students necessarily know the Verlet method?  no]
the Verlet method simplifies to 
the Euler method,
which is used to calculate the position at subsequent time steps.
%Most undergraduate students learn the Euler method in Computational Physics and Differential Equations, but don't encounter the Verlet method.
\begin{equation}
  \vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t).
    \label{eq:euler}
\end{equation}

The units of the simulated variables are summarized in Table ~\ref{tab:1}.

\begin{table}[h!]
\centering
\caption{Simulation units.}
\begin{ruledtabular}
\begin{tabular}{c c p{5cm}}
Quantity & Unit \\
\hline
length &  $a_0$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
dimensionless interaction strength & $q$ \\
effective colloidal charge & $Z^*$ \\
solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $E_0 / a_0$\\
viscosity/damping constant & $\eta$ \\
time &  $\eta / E_0$ \\
velocity &  $ E_0 / \eta a_0$ \\
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}

\section{Results}
\label{sec:results}
 We demonstrate how a single particle (Sec.~\ref{sec:one})
 and many particles
 move (Sec.~\ref{sec:sync})
 in response to this applied force in a variety of environments.
 In Sec.~\ref{sec:kink} we set $F^{DC}$
 to zero and track the motion
 of a high density area of a particle chain
 (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.
%
\subsection{Single particle system}
% In subsection headings, only the first word is capitalized.
\label{sec:one}

We drive a single particle across a periodic landscape 
along the y-direction 
\begin{equation}
  \label{eq:landscape_potential}
  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
\end{equation}
with $N_p=20$ troughs in the landscape
and landscape period $\lambda = L/N_p = 1.825 a_0$.
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
This is illustrated in Fig.~\ref{fig:1}(a) where
the red (blue) regions show local maxima (minima).
The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:landscape}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,
The numerical implementation of the landscape 
is calculated with
\begin{equation}
  \label{eq:force}
  F_y(y) = A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where $A_{p} = 0.1$.

Because we are performing molecular dynamics
for a single particle on a smooth potential energy landscape,
we use a large simulation time step of 
$\Delta t = 0.01$.
In Sec.~\ref{sec:sync},
we decrease the time step
to treat multiple particle interactions.

We drive the particle with
$F^{AC}=0.2$, $F^{DC}=0.01$, and $f=0.01$ cycles per time unit.
When the superposition of $F^{AC}$ and $F^{DC}$ is large enough
to overcome the barrier height of the landscape troughs,
the particle hops between troughs in the energy landscape.
In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position.
In Fig.~\ref{fig:0}(a)
we plot the applied driving force
as a function of time
and in Fig.~\ref{fig:0}(b) 
we show the position of the particle
as a function of time.
The initial position is near $y=0$
and the initial driving force is minimum
$F_D(t=0) = F^{DC}$.
Over the time period $T/2 = 50$ %time units
the particle moves
in the positive y-direction
while $F_{D}(t) > 0 $
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the period $T = 1/f = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the average displacement during one time period is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$. %, where $T = 1/f$.

%single DC case with figure
%dc_current 0.01
%ac_current 0.2
%ac_frequency 0.0001
%fp_y 0.1
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{position.png}
\caption{The position as a function of time of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

Phase locked steps occur for average displacements integer
multiples of the substrate period $n\lambda$.

We sample the average velocity $\langle v_y \rangle $
as a function of $F^{DC}$.
